A generalization of the Bradley-Terry model for draws in chess

### 2019-05-09 inst/collusion_chess_draws.Rnw
analyse curacao candidates match by match rather than using the table (which does not say who played white and who played black)

@@

@@

\section{Chess collusion}

@@

\begin{table}
\centering
\begin{tabular}{l|cccccc}
& \rotatebox{-90}{Fischer} 
&\rotatebox{-90}{Geller} 
&\rotatebox{-90}{Petrosian\ }
&\rotatebox{-90}{Korchnoi} 
&\rotatebox{-90}{Filip}
& $\cdots$
\\ \hline %       Fischer   Geller    Petrosian Korchnoi    Filip
Fischer     (USA)  &   -     & \halfb  & \halfw  & \oneww    & \halfb   &\\
Geller    (USSR)  & \halfw  &    -    & \halfb  & \halfw    & \halfw   &\\
Petrosian (USSR)  & \halfb  & \halfw  &    -    & \halfb    & \halfw   &\\
Korchnoi  (USSR)  & \zerob  & \halfb  & \halfw  &   -       & \oneww   &\\
Filip      (TCH)  & \halfw  & \halfb  & \halfb  & \zerob    &  -       &\\
$\vdots$&&&&&\\
\end{tabular}\label{interzonal}
\caption{Subset of the Fifth Interzonal round-robin, Stockholm 1962.
  Results (0,\halfw,1) and box colour refer to row player.  Thus, for
  example, Fischer played black against Geller and drew}
\end{table}

@@

In figure~\ref{collusion_fig}, note the high estimate for $D_s$ (also
note Fischer's dominance).  The maximum likelihood estimate is
$D\simeq 0.043, D_s=0.354$; a constrained maximization gives a
likelihood ratio of about 5096, or a $p$-value of about $3.6\times
10^{-5}$.  Indeed we could go further and plot a profile likelihood
curve for the log-odds ratio $\ln\left(D_s/D\right)$ which, if
Dirichlet, would be approximately Gaussian CITEKENDALANDSTUARTHERE.
This is illustrated in Figure~\ref{DOESNOTEXIST}, that shows a
credible interval of about HAVENOTDONETHISET.

@@

\begin{figure}[htbp]
\begin{center}
<<proflike,echo=TRUE,fig=TRUE>>=
plot(x,jj-max(jj),type='b',xlab="log-odds",ylab="support")
abline(h=c(0,-2))
@
\caption{Profile likelihood \label{proflike} for odds ratio
  $\log\left(D_s/D\right)$}
  \end{center}
\end{figure}

\bibliography{hyper2}
\end{document}

>>>>

\cite{moul2009}

\cite{davidson1970}

\cite{joe1990}

\cite{moul2009}: for instance, a player may have the choice of more
and less conservative strategies with different chances for a
draw. Risk- averse players may pursue less aggressive strategies,
trading expected payout for reduced variance

\cite{moul2009}: the question of Soviet collusion during the period
when the USSR dominated the struggle for the world championship has
been one of the long-standing debates in the chess world. The
best-known (but not unique) criticism was aired by Bobby Fischer some
weeks after the 1962 Candidates Tournament in Curacao

@@

\begin{equation}
\prod_{i<j}
\left(\frac{p_D}{p_i+p_j+p_D}\right)^{D_{ij}^\mathrm{white}+D_{ij}^\mathrm{black}}
\end{equation}

@@

\section{Chess collusion: Cura\c{c}ao 1962}
\begin{table}
\centering
\begin{tabular}{ll|cccccccc}
 &
&\rotatebox{-90}{Petrosian} 
&\rotatebox{-90}{Keres} 
&\rotatebox{-90}{Geller}
&\rotatebox{-90}{Fischer} 
&\rotatebox{-90}{Korchnoi}
&\rotatebox{-90}{Benko}
&\rotatebox{-90}{Tal}
&\rotatebox{-90}{Filip}\\
Petrosian &USSR    &             &DDDD    &DDDD   &DWDD   &DDWW   &DDWD   &WWD    &DWWD\\
Keres     &USSR    &             &NA      &DDDD   &LDWD   &DDWD   &WWWL   &WDW    &DWWD\\
Geller    &USSR    &             &NA      &NA     &WWDL   &DDWD   &DDDW   &DWW    &DWWD\\
Fischer   &USA     &             &NA      &NA     &NA     &LWLD   &LWDW   &DWD    &WDWD\\
Korchnoi  &USSR    &             &NA      &NA     &NA     &NA     &DDDL   &WLD    &WWWW\\
Benko     &USA     &             &NA      &NA     &NA     &NA     &NA     &WLD    &LWWD\\
Tal       &USSR    &             &NA      &NA     &NA     &NA     &NA     &NA     &WLD \\
Filip     &TCH     &             &NA      &NA     &NA     &NA     &NA     &NA     &A \\
\end{tabular}
\caption{Curacao 1962 SPELLED WRONG}
\end{table}

\section{Chess collusion: interzonal Stockholm 1962}

@@

\begin{table}
\centering
\begin{tabular}{l|cccccc}
&\rotatebox{-90}{Fischer} 
&\rotatebox{-90}{Geller} 
&\rotatebox{-90}{Petrosian\ }
&\rotatebox{-90}{Korchnoi} 
&\rotatebox{-90}{Filip}
& $\cdots$
\\ \hline %       Fischer   Geller    Petrosian Korchnoi    Filip
Fischer    (USA)  &   -     & \halfb  & \halfw  & \oneww    &\halfb    &\\
Geller    (USSR)  & \halfw  &    -    & \halfb  & \halfw    & \halfw   &\\
Petrosian (USSR)  & \halfb  & \halfw  &    -    & \halfb    & \halfw   &\\
Korchnoi  (USSR)  & \zerob  & \halfb  & \halfw  &   -       & \oneww   &\\
Filip      (TCH)  & \halfw  & \halfb  & \halfb  & \zerob    &  -       &\\
$\vdots$&&&&&\\
\end{tabular}\label{interzonal}
\caption{Subset of the Fifth Interzonal round-robin, Stockholm 1962.
  Results (0,\halfw,1) and box colour refer to row player.  Thus, for
  example, Fischer played black against Geller and drew}
\end{table}

@@

In figure~\ref{collusion_fig}, note the high estimate for $D_s$ (also
note
Fischer's dominance).

The maximum likelihood estimate is $D\simeq 0.043, D_s=0.354$; a constrained maximization gives a likelihood ratio of about 5096, or a $p$-value of about $3.6\times 10^{-5}$.  Indeed we could go further and plot a profile likelihood curve for the log-odds ratio $\ln\left(D_s/D\right)$ which, if Dirichlet, would be accurately parabolic over a wide range (OHAGAN AND FORSTER, p343).  This is illustrated in Figure~\ref{DOESNOTEXIST}, that shows a credible interval of about HAVENOTDONETHISET.

@@

\begin{figure}[htbp]
\begin{center}
<<proflike,echo=TRUE,fig=TRUE>>=
plot(x,jj-max(jj),type='b',xlab="log-odds",ylab="support")
abline(h=c(0,-2))
@
\caption{Profile likelihood \label{proflike} for odds ratio
  $\log\left(D_s/D\right)$: note the accurately parabolic figure in
  the region plotted}
  \end{center}
\end{figure}

\bibliography{chess}
\end{document}
